### 2017-03-16 articles/chemical-tagging/ms.tex
Fix Travis

%% This document is part of the snob project
%% Copyright 2017 the authors. All rights reserved.

\documentclass{aastex61}

% For revision history
\IfFileExists{vc.tex}{\input{vc.tex}}{
    \newcommand{\githash}{UNKNOWN}
    \newcommand{\giturl}{UNKNOWN}}

% Define commands
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}

\received{}
\revised{}
\accepted{}

\submitjournal{AAS Journals}

\shorttitle{Chemical tagging can work, but does it?}
\shortauthors{Casey et al.}

\begin{document}

\title{Chemical tagging can work, but does it?}

\correspondingauthor{Andrew R. Casey}
\email{andrew.casey@monash.edu}

\author[0000-0003-0174-0564]{Andrew R. Casey}
\affil{Monash Centre for Astrophysics, Monash University, Melbourne, 
	   VIC 3800, Australia}

\author{(some order of:}

\author[0000-0003-2952-859X]{John Lattanzio}
\affil{Monash Centre for Astrophysics, Monash University, Melbourne, 
	   VIC 3800, Australia}

\author[0000-0002-0583-5918]{David Dowe}

\author[0000-0002-1716-690X]{Aldeida Aleti}

\author{)}

\author{others?}

\begin{abstract}
Snobbery.
\end{abstract}

\keywords{}

\section{Introduction} 
\label{sec:introduction}

\acknowledgments
Wallace

\software{astropy \citep{Robitaille:2013}, numpy, scipy, scikit-learn}

\begin{thebibliography}{}

\bibitem[Astropy Collaboration et al.(2013)]{Robitaille:2013} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, A33

\end{thebibliography}

\end{document}