### 2017-09-20 articles/gmm-search/ms.tex
David's search

%% This document is part of the snob project
%% Copyright 2017 the authors. All rights reserved.
\documentclass{elsarticle}
\journal{Journal of Artificial Intelligence}
\usepackage{bm,url}
%\usepackage{boldsymbol}
\usepackage{verbatim} % for \begin{comment}'ing out sections.

% For revision history
\IfFileExists{vc.tex}{\input{vc.tex}}{
	\newcommand{\githash}{UNKNOWN}
	\newcommand{\giturl}{UNKNOWN}}

% Define commands
\newcommand{\article}{\emph{Article}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}

% Surveys
\newcommand{\apogee}{\acronym{APOGEE}}
\newcommand{\ges}{\acronym{GES}}
\newcommand{\hermes}{\acronym{HERMES}}
\newcommand{\galah}{\acronym{GALAH}}
\newcommand{\fourmost}{\acronym{4MOST}}
\newcommand{\weave}{\acronym{WEAVE}}
\newcommand{\gaia}{\project{Gaia}}

% Common terms
\newcommand{\vect}[1]{\boldsymbol{\mathbf{#1}}}

\def\teff{T_{\rm eff}}
\def\logg{\log{g}}

% Have some maaaath.
\def\infinity{\rotatebox{90}{8}}
\def\veccov{\vect{C}}
\def\vecmean{\vect{\mu}}
\def\vectheta{\vect{\theta}}
\def\weight{w}
\def\weights{\vect{\weight}}
\def\datum{y}
\def\data{\vect{\datum}}
\def\likelihood{\mathcal{L}}

% Affiliation(s)
\newcommand{\moca}{
	\affil{School of Physics and Astronomy, Monash University, 
		Melbourne, Clayton VIC 3800, Australia}}
\newcommand{\claytonfit}{
	\affil{Faculty of Information Technology, Monash University,
		Melbourne, Clayton VIC 3800, Australia}}
\newcommand{\caulfieldfit}{
	\affil{Faculty of Information Technology, Monash University,
		Melbourne, Caulfield East VIC 3145, Australia}}

\begin{document}
	\begin{frontmatter}
		\title{A search method for Gaussian mixtures using MML}

		%\author{Aleti, Casey, Dowe, Lattanzio}

		\begin{abstract}

		\end{abstract}

		\begin{keyword}
			Search\sep mml
		\end{keyword}
	\end{frontmatter}
\section{Introduction} 
\label{sec:introduction}

\section{MML}
\
There are $N$ data points each with $D$ dimensions, which are to be modelled
by $K$ mixture of $D$-dimensional gaussian distributions, each with a relative
weighting $\weight_k$ such that $\sum_{k=1}^{K}\weight_k = 1$. 
The data have the same error value, $y_{err}$, in all $D$ dimensions, for all $N$ observations.

The full expression for the message length is
given by,

\begin{eqnarray}
I_K & = & K\log{2} % I_k
    + \frac{(K - 1)}{2}\log{N} - \frac{1}{2}\sum_{k=1}^{K}\log{w_k} - \log{|\Gamma(K)|} % I_w
    + \mathcal{L}(\data|\vectheta) - DN\log{y_{err}} \\ % Likelihood \\
  & + & \frac{1}{2}\sum_{k=1}^{K}\left[\frac{D(D+3)}{2}\log{{Nw_k}} - (D + 2)\log{|\veccov_k|} - D\log{2}\right] % I_t
    - \frac{Q}{2}\log(2\pi) + \frac{\log{Q\pi}}{2} % lattice, minus the lattice of part 2
\end{eqnarray}

\noindent{}where $Q = \frac{1}{2}DK(D + 3) + K - 1$, the number of free parameters, and
$\mathcal{L}$ is the log-likelihood of a multivariate gaussian distribution.

Say we wanted to calculate whether another mixture was warranted. If another
mixture were preferred then we would want:

\begin{eqnarray}
  \Delta{}I_{K+1} - I_{K} < 0
\end{eqnarray}

The expression is given as:
\begin{eqnarray}
\Delta{I_{K+1} - I_K} & = & (K + 1)\log{2} - K\log{2} \\ % I_k^(new) - %I_k^(old)
  &&+ \frac{(K)}{2}\log{N} - \frac{1}{2}\sum_{k=1}^{K+1}\log{w_k}^{(new)} - \log{|\Gamma(K+1)|} \\ % I_w^(new)
  &&- \frac{(K - 1)}{2}\log{N} + \frac{1}{2}\sum_{k=1}^{K}\log{w_k} + \log{|\Gamma(K)|}\\ % I_w^(old)
  &&+ \mathcal{L}^{(new)} - DN\log{y_{err}} \\ % Likelihood (new)\\
  &&- \mathcal{L}^{(old)} + DN\log{y_{err}} \\ % Likelihood (old) \\
  &&+ \frac{1}{2}\sum_{k=1}^{K+1(new)}\left[\frac{D(D+3)}{2}\log{{Nw_k}} - (D + 2)\log{|\veccov_k|} - D\log{2}\right] \\ % I_t^(new)
  &&- \frac{1}{2}\sum_{k=1}^{K(old)}\left[\frac{D(D+3)}{2}\log{{Nw_k}} - (D + 2)\log{|\veccov_k|} - D\log{2}\right] \\ % I_t^(old)
  &&- \frac{Q^{(new)}}{2}\log(2\pi) + \frac{\log{Q^{(new)}\pi}}{2} \\ % lattice, minus the lattice of part 2
  &&+ \frac{Q^{(old)}}{2}\log(2\pi) - \frac{\log{Q^{(old)}\pi}}{2} % lattice, minus the lattice of part 2
\end{eqnarray}

\noindent{}By making use of $\log{\Gamma(K)} - \log{\Gamma(K + 1)} = -\log{K}$ and re-arranging the expression:

\begin{eqnarray}
\Delta{}I_{K+1} - I_K &=& \log{2} % \Delta{}I_k
    + \frac{1}{2}\log{N} - \log{K} - \frac{1}{2}\left(\sum_{k=1}^{K+1}\log{w_k^{(new)}} - \sum_{k=1}^{K}\log{w_k^{(old)}}\right) \nonumber \\ % \Delta{}I_w
&& + \mathcal{L}^{(new)} - \mathcal{L}^{(old)} \nonumber \\ % likelihood
&& + \frac{1}{2}\left[\frac{D(D+3)}{2}\left(\sum_{k=1}^{K+1}\log{Nw_k^{(new)} - \sum_{k=1}^{K+1}\log{Nw_k^{(old)}}} \right) - \left(D+2\right)\left(\sum_{k=1}^{K+1}\log{|\veccov_k|^{(new)}} - \sum_{k=1}^{K+1}\log{|\veccov_k|^{(old)}}\right)\right] \nonumber \\
&& + \frac{\log(2\pi)}{2}(Q^{(old)} - Q^{(new)}) + \frac{\pi}{2}\left(\log{Q^{(new)}} - \log{Q^{(old)}}\right)
\label{eq:13}
\end{eqnarray}

Expanding the $Q$ terms:

\begin{eqnarray}
Q^{(old)} - Q^{(new)} &=& \frac{1}{2}DK(D + 3) + K - 1 - \frac{1}{2}D(K + 1)(D + 3) + (K + 1) - 1 \nonumber \\
Q^{(old)} - Q^{(new)} &=& -\frac{1}{2}D(D+3) + 2K  - 1
\label{eq:14}
\end{eqnarray}

\noindent{}And making use of the following logarithmic identities,
\begin{eqnarray}
  \log{Q^{(new)}} &=& \log{\left(\frac{1}{2}D(K+1)(D + 3) + K\right)} \nonumber \\
                  &=& \log{\left(\frac{1}{2}D(K+1)(D + 3)\right)} + \log{\left(1 + \frac{K}{\frac{1}{2}D(K+1)(D + 3)}\right)} \\
  \log{Q^{(old)}} &=& \log{\left(\frac{1}{2}DK(D + 3) + K - 1\right)} \nonumber \\
                  &=& \log{\left(\frac{1}{2}DK(D + 3)\right)} + \log{\left(1 + \frac{K - 1}{\frac{1}{2}DK(D + 3)}\right)}
\end{eqnarray}

\noindent{}gives us,

\begin{eqnarray}
  \log{Q^{(new)}} - \log{Q^{(old)}} &=& \log{\left(\frac{1}{2}D(K+1)(D + 3)\right)} - \log{\left(\frac{1}{2}DK(D + 3)\right)} \nonumber \\
                                    &&+ \log{\left(1 + \frac{K}{\frac{1}{2}D(K+1)(D + 3)}\right)} - \log{\left(1 + \frac{K - 1}{\frac{1}{2}DK(D + 3)}\right)} \quad .\\
\end{eqnarray}
The second row of terms can be ignored because they are very small (typically less than 1 bit). This is because as $K \rightarrow \infinity$, $2K/D(K+1)(D+3) \rightarrow 1$, thus $\log{\left(1 + \frac{K}{\frac{1}{2}D(K+1)(D + 3)}\right)} \rightarrow \log{2}$. Similarly as $D \rightarrow \infinity$, $2K/D(K+1)(D+3) \rightarrow 0$.

As $K \rightarrow \infinity$ then $2(K-1)/DK(D+3) \rightarrow 1$ and as $D \rightarrow \infinity$ then $2(K-1)/DK(D+3) \rightarrow 0$ and thus $\log{\left(1 + \frac{K - 1}{\frac{1}{2}DK(D + 3)}\right)} \approx 0$.

\noindent{}Ignoring these minor terms:

\begin{eqnarray}
  \log{Q^{(new)}} - \log{Q^{(old)}} &\approx& \log{\left(\frac{1}{2}D(K+1)(D + 3)\right)} - \log{\left(\frac{1}{2}DK(D + 3)\right)} \nonumber \\
  \log{Q^{(new)}} - \log{Q^{(old)}} &\approx& \log{(K + 1)} - \log{K}
  \label{eq:19}
\end{eqnarray}

\noindent{}Substituting Eqs. \ref{eq:19} and \ref{eq:14} into \ref{eq:13} yields:

\begin{eqnarray}
\Delta{}I_{K+1} - I_K &\approx& \log{2} % \Delta{}I_k
    + \frac{1}{2}\log{N} - \log{K} - \frac{1}{2}\left(\sum_{k=1}^{K+1}\log{w_k^{(new)}} - \sum_{k=1}^{K}\log{w_k^{(old)}}\right) \nonumber \\ % \Delta{}I_w
&& + \mathcal{L}^{(new)} - \mathcal{L}^{(old)} \nonumber \\ % likelihood
&& + \frac{1}{2}\left[\frac{D(D+3)}{2}\left(\sum_{k=1}^{K+1}\log{Nw_k^{(new)} - \sum_{k=1}^{K+1}\log{Nw_k^{(old)}}} \right) - \left(D+2\right)\left(\sum_{k=1}^{K+1}\log{|\veccov_k|^{(new)}} - \sum_{k=1}^{K+1}\log{|\veccov_k|^{(old)}}\right)\right] \nonumber \\
&& + \frac{\log(2\pi)}{2}(-\frac{1}{2}D(D+3) + 2K  - 1) + \frac{\pi}{2}\left(\log{(K + 1)} - \log{K}\right)
\end{eqnarray}

\section{The k-means++ algorithm}

The k-means++ algorithm extends the k-means method with an efficient way of choosing centres. Lets $D(x)$ denote the shortest distance from a data point $x$ to the closest centre that has already been selected. The k-means++ performs the following steps:

\begin{enumerate}
	\item Selects a centre $c_1$ uniformly at random from the $N$ data points.
	\item Selects a new centre $c_i$ from the data points with probability $\frac{D(x)^2}{\sum_{x\in X}}D(x)^2$
	\item Repeat Step 2. until k centres have been selected.
	\item Proceed as with the standard k-means algorithm.
\end{enumerate}

k-means++ is O(log k)-Competitive.

\section{The Heuristic}

Re the case of two (or more) classes with the same mean:
In 2D, a simple case in point is two classes with the same mean
such that $sigma_{1, x} = sigma_{2, y} >> sigma_{2, x} = sigma_{1, y}$.
Andy said that this and higher dimensional stuff can be picked up
by dropping to 1 dimension along the eigenvector of largest eigenvalue.
I like that.
I wonder also about the merits of dropping to 2 dimensions (if we have
at least 2 attributes), being the eigenvectors of the two largest eigenvalues.
If it could be done quickly, one could transform to a circle, and then
some test (e.g., Kolmogorov-Smirnov test) for whether data is uniform
in radial distribution [0, 2 $\pi$).  If not uniform, then perhaps grounds for
splitting.

One could possibly check whether uniform radially directly from the
original 2D projection without having to project it to a circle (i.e., it
might be pretty easy to work out what the angle of each point would
be post-transformation without having to do the transformation).  The
2D projection is 1D in the sense that we only care about the angle
around the circle.  And, again, Kolmogorov-Smirnov is a popular test
for uniformity.

Re looking at things in 1 dimension (not 2D followed by 1D [0, 2 pi) angle
as above but rather projecting on eigenvector of largest eigenvalue), we
discussed moments - e.g., \url{https://en.wikipedia.org/wiki/Normal_distribution} .
So, 2nd, 4th, 6th and 8th moments of a true Gaussian are 1, 3, 15, 105.

The heuristic I'm going to propose is independent (i.e., I haven't heard
of it before), but it seems pretty easy, it mightn't be original, and I can
conceive that Andy might already be using it.

Let's suppose we have a 1D component.  We'll shift it (without loss of
generality, w.l.o.g.) so that its mean is 0 and we'll shrink or expand it
(again, w.l.o.g.) so that its s.d. ($\sigma$) is 1 (and variance v is 1).

We consider now splitting it into K components with mean 0,
mixing weights (or proportions) $p_1, ..., p_i, ..., p_K$ and s.d.s $\sigma_i$ (and variances $v_i = (\sigma_i)^2$).
Let's set $K = 2, so p_2 = 1 - p_1$.

Suppose the moments of our current component are 1, $3 \alpha_4, 15 \alpha_6 and 105 \alpha_8, ... $.
We get some equations - if we have the correct number of equations,
they're simultaneous equations.

If we get too many equations (by using too many moments), then it
becomes something of a fitting or regression problem to choose
$p_1, ..., p_i, ..., p_K$ and s.d.s $\sigma_i$, variances $v_i = (\sigma_i)^2$.

Our simultaneous equations are:
$p_1 v_1       + (1 - p_1) v_2       =    1$
$p_1 (v_1)^2  + (1 - p_1) (v_2)^2  =    3 \alpha_4$
$p_1 (v_1)^3  + (1 - p_1) (v_2)^3  =   15 \alpha_6$
$p_1 (v_1)^4  + (1 - p_1) (v_2)^4  =  105 \alpha_8$
(and possibly etc., as possibly needed).

These might be messy to solve on the fly  but
one could pre-process by solving many of these beforehand
(actually, approximately solving these beforehand, as it's only
a heuristic, anyway) at the very start of the program while
loading in the data.

(I remember that Chris Wallace got me to do this a similar pre-processing step when dealing with von Mises angular/circular data.)
So, for $1000 x 1000 = 10^6$ different values of $\alpha_4 and \alpha_6$
or for $100 x 100 x 100 = 10^6$ different values of $\alpha_4, \alpha_6 and \alpha_8$
one could pre-compute {\em approximate} (rough heuristic) 1,000,000ish or so `solutions' to the above simultaneous equations.

Then, when running the program with real actual values of $\alpha_4$ and $\alpha_6$ (and $\alpha_8$),
one could just grab something from nearby in this pre-computed look-up table
as putative values for $p_1, v_1, v_2$ (and etc., if need be).

And, of course, one could change K from 2 to something bigger.

There should still be some use of randomness in the search.

\begin{thebibliography}{}

% Note: I include the DOI as a comment on the same line after every \bibitem
%       entry. I do this in case I ever need to extract full bibstems for the
%       bibliography, which can be done using this curl command and the DOI:

%       curl -LH "Accept: application/x-bibtex" http://dx.doi.org/10.5555/12345678

%       If no DOI is available then I list the ADS identifier.
\end{thebibliography}

\end{document}